BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-07-17 Msc-Slides/main.tex
Existing changed with available on slide 7
Slides re-worked once again

\begin{frame}
    \frametitle{Malware: an increasingly significant problem }

@@

    Two main types:
    \medskip
    \begin{itemize}
        \item \textcolor{sapienza}{\textbf{Static Analysis:}}\\
        involves the inspection of the different data and code sections of a binary
        \item \textcolor{sapienza}{\textbf{Dynamic Analysis:}}\\
        the malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{0.5cm}

         \begin{beamerboxesrounded}[shadow=true]{}
    Dynamic analysis strongly favoured as it allows to dodge \\ code obfuscations and deal with a large number of samples
    \end{beamerboxesrounded}

    \begin{textblock*}{2cm}(9.0cm,-6.1cm)
   \includegraphics[width=1.2cm]{image/search.png}% use the \includegraphics command here
	\end{textblock*}

@@

\begin{frame}[fragile]
    \frametitle{Implementation of function call monitoring}

@@

One technique is \textbf{Dynamic Binary Instrumentation (DBI)}.
\begin{itemize}
\item[$\rightarrow$] The behaviour of an application is inspected at run-time, without the need of recompiling it, via the injection of analysis code
\end{itemize}
\smallskip

@@

\subsection{Evasion}
\begin{frame}
    \frametitle{The threat posed by evasive malware}

@@

    \begin{figure}
    	\vspace{-0.6cm}
    	\hspace*{0.5cm}
        \includegraphics[width=11cm]{image/evasive.pdf}
    \end{figure}

@@

\begin{frame}
    \frametitle{Existing problems}

\section{Problem}	
Two main problems which need to be addressed:
\\\bigskip
\begin{enumerate}
\item Existing API tracing tools have \textbf{limited logging capabilities}
\smallskip
\begin{itemize}
\item[--] The amount of recorded information is too little
\item[--] It is hard to distinguish the calls made directly from the sample's main executable from the ones made by libraries
\end{itemize}
\bigskip
\item Current API hooking techniques are \textbf{easily detectable} and are not coupled with mechanisms to hide their presence
from evasive malware
\end{enumerate}

@@

\subsection{Overview}
\begin{frame}
    \frametitle{BlueTracer}
    \setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

\begin{beamerboxesrounded}[shadow=true]{}
\textbf{BlueTracer} is a robust library and system call tracer for Windows programs specialized in evasive malware
\end{beamerboxesrounded}

\textcolor{sapienza}{\textbf{The tool possesses a remarkable logging power:}}
\begin{itemize}
\item Access to a source of calls related information is required
\item Logging operations are made challenging by the heterogeneity of Windows libraries used in malware and the lack of well-structured documentation for their prototypes
\begin{itemize}
\item[$\rightarrow$] Integration of reliable external sources \\ (\textbf{Dr. Memory}
and \textbf{CISCO PyREBox})
\end{itemize} 
\end{itemize}

\textcolor{sapienza}{\textbf{Key features:}}
\begin{itemize}
\item Undetected tracing of input parameters, output buffers and return values of over 17 000 system calls and library calls
\item Logging of asynchronous events
\item Allows to trace only calls from the main executable
\end{itemize}

\begin{frame}
    \frametitle{BlueTracer (cont'd)}
\setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

\textcolor{sapienza}{\textbf{Solution to the detection problem:}}
\begin{itemize}
\item Seamless integration with \textbf{BluePill}, a stealthy execution framework based on \textbf{Intel Pin}
%\item Combines reliable external sources of prototypes information
\end{itemize}

\bigskip
	    \begin{figure}
    	\vspace{-0.3cm}
    	%\hspace*{1cm}
        \includegraphics[width=6cm]{image/BluePill.pdf}
    \end{figure}

\textcolor{sapienza}{\textbf{Faced implementation challenges:}}
\begin{itemize}
\item Making best use of
Intel Pin's capabilities with respect to run-time CPU and memory costs
\item Addressing some inherent limitations of Intel Pin

\end{itemize}

@@

\section{Evaluation}

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Evaluation with Al-Khaser}
	\textbf{Al-Khaser} is an open-source application which performs common checks employed by malware families to determine if they are being
executed in an analysis environment.
\\\medskip
Checks divided in categories:
\begin{itemize}
\item \textbf{Anti-Debugging}
\item \textbf{Timing-based}
\item \textbf{Human Interaction Detection}
\item \textbf{Anti-Virtualization}
\item \textbf{Anti-Analysis}
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
BlueTracer: 
\begin{itemize}
\item remained undetected thanks to its integration with BluePill
\item managed to track all the checks
\end{itemize}
\end{beamerboxesrounded}

@@

\subsection{Evasive malware samples}
\begin{frame}
    \frametitle{Evaluation with evasive malware samples}

@@

	\textbf{\textcolor{sapienza}{Contribution:}} \\
	\smallskip	
	\tab Design and implementation of \textbf{BlueTracer},
a robust library and system call tracer for Windows programs specialized in evasive
malware.
	\\\bigskip
	\textbf{\textcolor{sapienza}{Future Developments:}} \\
	\begin{itemize}
	\item Automatic methodology for large-scale evaluation 
	\item Improvement of logging capabilities
	\item Usage of log filtering techniques
	\end{itemize}

>>>>

\iffalse
\begin{frame}
    \frametitle{Malware: an increasingly significant problem }

@@

	Malware is an ever-growing threat

    %Two main types:
    \medskip
    \begin{itemize}
        \item \textcolor{sapienza}{\textbf{Static Analysis:}}\\
        involves the inspection of the different data and code sections of a binary
        \item \textcolor{sapienza}{\textbf{Dynamic Analysis:}}\\
        the malware sample is executed and the actions it performs on the environment are observed
    \end{itemize}
    \vspace{0.5cm}

         \begin{beamerboxesrounded}[shadow=true]{}
    Dynamic analysis strongly favoured as it allows to dodge most code obfuscations and deal with a large number of samples
    \end{beamerboxesrounded}

    \begin{textblock*}{2cm}(8.0cm,-6.1cm)
   \includegraphics[width=1.2cm]{image/search.png}% use the \includegraphics command here
	\end{textblock*}

@@

\iffalse
\begin{frame}[fragile]
    \frametitle{Implementing function call monitoring}

@@

One technique is \textbf{Dynamic Binary Instrumentation (DBI)}.
\begin{itemize}
\item[$\rightarrow$] The behaviour of an application is inspected at run time, without the need of recompiling it - by injecting analysis code
\end{itemize}
\smallskip

@@

\subsection{Problems}
\begin{frame}[fragile]
    \frametitle{Problem 1: limited logging capabilities}

Available API tracing tools have \textbf{limited logging capabilities}
%\begin{itemize}
%\item[--] The amount of recorded information is too little
%\end{itemize}
\medskip
	\begin{block}{}
\begin{lstlisting}[basicstyle=\ttfamily\large,xleftmargin=50pt]
<@\textcolor{red}{record\_before(libcall\_name, arg1)}@> 
retval = libcall(arg1, &arg2)
<@\textcolor{red}{record\_after(retval, *arg2)}@> 
\end{lstlisting}
\end{block}	
\medskip
\begin{itemize}
\item Access to a source of calls related information is required
\begin{itemize}
\item[--] Prototype (number of arguments, data types, input/output)
%\item[--] Extraction
\end{itemize}
\medskip
\item \textbf{Challenge:} heterogeneity of Windows libraries used in malware and lack of well-structured documentation for their prototypes
\end{itemize}

\begin{frame}[fragile]
    \frametitle{Problem 2: logging only calls made by sample}

It is hard to distinguish the calls made directly by the sample from the ones made within libraries
\begin{itemize}
\item Resulting logs are large and full of irrelevant information
\end{itemize}

    \begin{figure}
    	\vspace{-0.2cm}
    	\hspace*{-1cm}
        \includegraphics[width=12cm]{image/MainImage1.pdf}
    \end{figure}

\begin{frame}[fragile]
    \frametitle{Problem 3: evasive malware}

     \begin{beamerboxesrounded}[shadow=true]{Evasive malware}
Malware that conceals its harmful behaviour when detecting a
hostile environment, such as a well-known sandbox
solution
    \end{beamerboxesrounded}
    \bigskip
    \bigskip
    \begin{figure}
    	\vspace{-0.8cm}
    	\hspace*{0.5cm}
        \includegraphics[width=10cm]{image/evasive.pdf}
    \end{figure}
\vspace{-0.3cm}
Current tracing techniques are \textbf{easily detectable} and are not coupled with mechanisms to hide their presence

\iffalse
\subsection{Evasion}
\begin{frame}
    \frametitle{The threat posed by evasive malware}

@@

    \begin{figure}
    	\vspace{-0.6cm}
    	\hspace*{0.5cm}
        \includegraphics[width=12cm]{image/evasive.pdf}
    \end{figure}

\iffalse
\begin{frame}
    \frametitle{Existing problems}

\section{Problem}	
Two main problems to address:
\\\bigskip
\begin{enumerate}
\item Available API tracing tools have \textbf{limited logging capabilities}
\smallskip
\begin{itemize}
\item[--] The amount of recorded information is too little
\item[--] It is hard to distinguish the calls made directly by the sample from the ones made within libraries
\end{itemize}
\bigskip
\item Current API hooking techniques are \textbf{easily detectable} and are not coupled with mechanisms to hide their presence
from evasive malware
\end{enumerate}

\subsection{Overview}
\begin{frame}
    \frametitle{BlueTracer: accurate API logging}
    \setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

\begin{beamerboxesrounded}[shadow=true]{}
\textbf{BlueTracer} is a robust library and system call tracer for Windows programs specialized in evasive malware based on Intel Pin
\end{beamerboxesrounded}

\bigskip

\textcolor{sapienza}{\textbf{The tool possesses a remarkable logging power:}}
\begin{itemize}
%\item Integration of reliable external sources \\ (\textbf{Dr. %Memory}
%and \textbf{CISCO PyREBox})
%\item Logging operations are challenging 
%by the heterogeneity of Windows libraries used in malware and the lack of well-structured documentation for their prototypes
%\begin{itemize}
\item Integration of reliable external sources \\ (\textbf{Dr. Memory}
and \textbf{CISCO PyREBox})
%\end{itemize} 
%\end{itemize}

%\textcolor{sapienza}{\textbf{Key features:}}
%\begin{itemize}
\item Stealthy tracing of input parameters, output buffers and return values of over 17,000 system calls and library calls
\item Logging of asynchronous events
%\item Allows to trace only calls from the main executable
\end{itemize}

@@

\begin{frame}
    \frametitle{BlueTracer: robust against evasive malware}
\setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

\textcolor{sapienza}{\textbf{Solution to the detection problem:}}
\begin{itemize}
\item Seamless integration with the \textbf{BluePill} stealthy execution framework
%\item Combines reliable external sources of prototypes information
\end{itemize}

\bigskip
\bigskip
	    \begin{figure}
    	\vspace{-0.8cm}
    	%\hspace*{1cm}
        \includegraphics[width=8cm]{image/BluePill.pdf}
    \end{figure}

%\textcolor{sapienza}{\textbf{Faced implementation challenges:}}
%\begin{itemize}
%\item Optimal use of
%Intel Pin's capabilities with respect to run-time CPU and memory costs
%\item Addressing some inherent limitations of Intel Pin
%\end{itemize}

@@

\begin{frame}
    \frametitle{BlueTracer: focussed tracing of the sample's actions}
\setbeamerfont*{itemize/enumerate subbody}{parent=itemize/enumerate body}

Only calls made directly by the sample are recorded through the use of context-sensitive introspection
\begin{itemize}
\item Stack pointer inspection
\end{itemize}    
\medskip
	    \begin{figure}
    	\vspace{-0.3cm}
    	%\hspace*{1cm}
        \includegraphics[width=10cm]{image/MainImageTree.pdf}
    \end{figure}

@@

\section{Validation}

\subsection{Al-Khaser}
\begin{frame}
    \frametitle{Validation (1) with Al-Khaser}
    \begin{beamerboxesrounded}[shadow=true]{}
	\textbf{Al-Khaser} is an open-source application which performs common checks employed by malware families to determine if they are being
executed in an analysis environment.
\end{beamerboxesrounded}
\medskip
Checks divided in categories:
\begin{itemize}
\item \textbf{Anti-Debugging}
\item \textbf{Timing-based}
\item \textbf{Human Interaction Detection}
\item \textbf{Anti-Virtualization}
\item \textbf{Detection of Analysis Tools}
\end{itemize}
\medskip
\begin{beamerboxesrounded}[shadow=true]{}
BlueTracer: 
\begin{itemize}
\item remained undetected thanks to its integration with BluePill
\item managed to track all the checks
\end{itemize}
\end{beamerboxesrounded}

@@

\subsection{Evasive malware samples}
\begin{frame}
    \frametitle{Validation (2) with highly evasive malware samples}

@@

	\textbf{\textcolor{sapienza}{Contribution:}} \\
	\smallskip	
	\tab Design and implementation of \textbf{BlueTracer},
a robust library and system call tracer for Windows programs specialized in evasive
malware.
	\\\bigskip
	\textbf{\textcolor{sapienza}{Future Developments:}} \\
	\begin{itemize}
	\item Automatic methodology for large-scale evaluation 
	\item Improvement of logging capabilities
	\item Usage of log filtering and aggregation techniques
	\end{itemize}

### 2018-07-17 conclusions.tex
Slides re-worked once again

Another future direction regarding BlueTracer might be the application of Big Data algorithms to the logs it produces. In fact, due to the wealth of information being recorded, the logs produced by BlueTracer are often very large (up to tens of gigabytes), making it challenging to extract valuable insights from them. In the specific case of BlueTracer, the information of interest is represented by the set of malicious and anti-analysis actions performed by malware. As a result of this, it might be a good idea to applying Big Data analytics to isolate just the relevant information and to even compare and integrate the data coming from different logs, thus making the analyst's life easier.

>>>>

Another future direction regarding BlueTracer might be the application of Big Data algorithms to the logs it produces. In fact, due to the wealth of information being recorded, the logs produced by BlueTracer are often very large (up to tens of gigabytes), making it challenging to extract valuable insights from them. In the specific case of BlueTracer, the information of interest is represented by the set of malicious and anti-analysis actions performed by malware. As a result of this, it might be a good idea to apply Big Data analytics to isolate just the relevant information and to even compare and integrate the data coming from different logs.% thus making the analyst's life easier.
